### 2008-12-23 doc/sec/SecurityFrameworkofARC1.tex
Adding VOMS attributes description.

% This file was converted to LaTeX by Writer2LaTeX ver. 0.5

@@

\documentclass{article}

@@

\usepackage{amsmath,amssymb,amsfonts,textcomp}

@@

\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}

@@

\renewcommand\thesection{\arabic{section}.}
\renewcommand\thesubsection{\arabic{subsection}.}
\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}.}
\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}.}

@@

\newcommand\liststyleLi{%

@@

\setlength\paperwidth{8.2673in}
\setlength\paperheight{11.6925in}

@@

@@

\bigskip
\clearpage\clearpage\setcounter{page}{1}\pagestyle{Convertv}

@@

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{delegtion.pdp} -- Delegation PDP}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The Delegation PDP is basically similar to Arc PDP, except it uses the
delegation policy parsed from remote proxy credential by
\textit{delegation.collector}, and evaluates the request against
delegation policy. See section 6for the design idea and use case of
delegation policy in fine-grained identity delegation.}

@@

{\selectlanguage{english}\itshape\color{black}
 Rule (1-)}

{\selectlanguage{english}\itshape\color{black}
  Subjects (1)}

{\selectlanguage{english}\itshape\color{black}
   Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
    Attribute (1-)}

{\selectlanguage{english}\itshape\color{black}
   Resources (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Resource (1-)}

{\selectlanguage{english}\itshape\color{black}
  Actions (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Action (1-)}

{\selectlanguage{english}\itshape\color{black}
  Conditions (0-1)}

{\selectlanguage{english}\itshape\color{black}
   Condition (1-)}

{\selectlanguage{english}\itshape\color{black}
    Attribute (1-)}

@@

{\selectlanguage{english}\itshape\color{black}
 RequestItem (1-)}

{\selectlanguage{english}\itshape\color{black}
  Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
   SubjectAttribute (1-)}

{\selectlanguage{english}\itshape\color{black}
  Resource (0-)}

{\selectlanguage{english}\itshape\color{black}
  Action (0-)}

{\selectlanguage{english}\itshape\color{black}
  Context (0-)}

{\selectlanguage{english}\itshape\color{black}
   ContextAttribute (1-)}

@@

\subsubsection{Basic Elements of Policy}
{\selectlanguage{english}\upshape\color{black}
There are 2 basic objects - {\textquotedbl}policy{\textquotedbl} and
{\textquotedbl}request{\textquotedbl}. There is 1 main actor -
Evaluator. Curretly there are two types of elements in policy: Policy
and Rule. Policy element is made of Rule elements.}

{\color{black}
Evaluator matches request to policy and produces one of 4 following
results:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{PERMIT} - policy explicitely permits activity specified in
request because request matches some part of policy and corresponding
effect specified in policy is PERMIT. }
\end{itemize}
{\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: }}\textit{PERMIT
person ALICE to PLAY in place called WONDERLAND}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request: }}\textit{person
ALICE wants to PLAY in place called WONDERLAND}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{DENY} - policy explicitely denies activity specified in Request
because Request matches some part of policy and corresponding effect
specified in policy is DENY.}
\end{itemize}
{\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Rule: DENY }}\textit{fruit
APPLE to GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{\ \ \ \ \ \ }}\textbf{\textit{Request}}\textit{: fruit
APPLE to be GROWN on PEACH tree}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{INDETERMINATE} - request has some part which does not correspond
to policy.}
\end{itemize}
{\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
be GROWN on WHEAT ground}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: flower SUNFLOWER
to be grown on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Explaination}}: Here, it is not possible to obtain any
matching result - neither positive (DENY or PERMIT) nor negative
(NOT\_APPLICABLE, see below)}

{\color{black}
In the request, the {\textquotedbl}ground{\textquotedbl} is completely
uncomparable to the {\textquotedbl}tree{\textquotedbl} in policy. One
can compare {\textquotedbl}PEACH tree{\textquotedbl} and
{\textquotedbl}APPLE tree{\textquotedbl} because they are both
{\textquotedbl}tree{\textquotedbl}; But it is impossible to compare
{\textquotedbl}PEACH tree{\textquotedbl} and {\textquotedbl}WHEAT
ground{\textquotedbl} becaue they of different kind (Policy is about
tree and Request is about ground).}

{\color{black}
In a similar way one can{\textquotesingle}t compare {\textquotedbl}fruit
APPLE{\textquotedbl} and {\textquotedbl}flower SUNFLOWER{\textquotedbl}
(here policy is about fruits and Request is about flower).}

{\color{black}
Any other situation which makes it impossibile to compare two attributes
will also cause {\textquotedbl}INDETERMINATE{\textquotedbl}.}

@@

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{NOT\_APPLICABLE} - all parts of the Request have corresponding
parts in the Policy, but some value of those parts are not the same.
Hence request does not match policy.}
\end{itemize}
{\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Rule}}\textit{: DENY fruit APPLE to
GROW on PEACH tree}}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit APPLE to
be GROWN on APPLE tree }}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
be GROWN on PEACH tree }}

{\selectlanguage{english}\upshape\color{black}
\textit{\ \ \ \ \ \ }\textbf{\textit{Request}}\textit{: fruit ORANGE to
be GROWN on APPLE tree }}

{\color{black}
Explanation: for each part of the Request evaluator can find relevant
part in the Policy - both Policy and Request are about fruit and tree.
But the values do not match.}

{\color{black}
If it is required to reduce evaluation results to boolean value PERMIT
maps to TRUE and rest of results to FALSE.}

{\selectlanguage{english}\upshape\color{black}
\textbf{\textit{Note}}\textit{: It would be useful to make it possible
to specify secondary effect which would become }active in case Request
is NOT\_APPLICABLE. For example:}

{\itshape\color{black}
DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT}

{\color{black}
But one should be careful because example above would allow fruit PLUMS
to grow on APPLE trees :)}

{\color{black}
This kind of requirement can be supported by using the algorithm between
policies. \ For example, in case of above scenario, we can use some
algorithm like {\textquotedbl}Permit-if-notapplicable{\textquotedbl}.
See below the {\textquotedbl}Policy matching{\textquotedbl} part for
more explaination. \ }

\subsubsection{Policy MAtching}
{\selectlanguage{english}\upshape\color{black}
Policy is made of Rule elements. Request is evaluated against each Rule.
Each evaluation produces same results as policy evaluation described
above. The results from all Rules are then combined in order to produce
final result for whole policy. Results Combining Algorithm is specified
in Policy. There are 26 algorithms currently:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textbf{\textit{Deny-Overrides}} - this is default if no algorithm
specified.}

\begin{itemize}
\item {\color{black}
If there is at least one DENY in results final result is DENY.}
\item {\color{black}
Otherwise if there is at least one PERMIT, the final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
\textbf{NOTE}\textbf{: CHECK if algorithm description is correct}}

{\color{black}
Special case is Policy wih no rules. Probably such policy shoudl be
treated as always producing DENY.}

{\bfseries\color{black}
Discussion}

{\itshape\color{black}
If there is no Rule under Policy, it means no restriction from this
Policy, what exactly shoud the Policy give? DENY or NOT\_APPLICABLE?}

{\itshape\color{black}
It seems to be more logical to produce INDETERMINATE because there is no
rule with elements which could be compared to request.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\bfseries\itshape\color{black}
Permit-Overrides.}

\begin{itemize}
\item {\color{black}
If there is at least one PERMIT in results final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\color{black}
Special case is Policy wih no rules. Probably such policy should be
treated as always producing DENY.}

{\bfseries\color{black}
NOTE: CHECK if algorithm description is correct}

{\bfseries\color{black}
Discussion}

{\itshape\color{black}
Question? The same as above.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\bfseries\itshape\color{black}
Ordered algorithms.}
\end{itemize}
{\selectlanguage{english}\color{black}
These specify priorities for all four possible results. Their names look
like Result1-Result2-Result3-Result4 with Result\# naming result types,
for example Permit-Deny-NotApplicable-Indeterminate. The results are
combined in following way:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item \begin{itemize}
\item {\color{black}
If there is at least one result of Result1 type then final result is
Result1.}
\item {\color{black}
Otherwise if there is at least one result of Result2 type then final
result is Result2.}
\item {\color{black}
Otherwise if there is at least one result of Result3 type then final
result is Result3.}
\item {\color{black}
Otherwise final result is Result4.}
\end{itemize}
\end{itemize}
{\color{black}
There are 24 possible combinations of those algorithms.}

{\selectlanguage{english}\upshape\color{black}
\textbf{Note:} It would be useful to have more combining algorithms. For
example }

@@

\begin{itemize}
\item {\color{black}
If this Rule gives DENY in results, final result is DENY.}
\item {\color{black}
Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.}
\item {\color{black}
Otherwise final result is INDETERMINATE. }
\end{itemize}
\item {\selectlanguage{english}\color{black}
{\textquotedblleft}Permit-if-allPermit{\textquotedbl} - Permit if all
the Rules gives Permit, this algorithm is useful in case if we are
collecting different policies from a few sources, and we want the
request to satisfy all of them.}

\begin{itemize}
\item {\color{black}
If all of the Rule give PERMIT, the final result is PERMIT.}
\item {\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\item {\color{black}
OnlyOneApplicable}

\begin{itemize}
\item {\color{black}
If there is one gives INDETERMINATE, final result INDETERMINATE is given
immediately.}
\item {\color{black}
Otherwise if there is exactly only one gives applicable result (DENY or
PERMIT), final result is as this result.}
\item {\color{black}
Otherwise if there is more than one gives applicable result, final
result is INDTERMINATE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
{\color{black}
This algorithm makes sure that only one Rule is selected when making
decision.}

\liststyleWWviiiNumxvii
\begin{itemize}
\item {\color{black}
FirstApplicable}

\begin{itemize}
\item {\color{black}
If there is one give DENY, PERMIT or INDETERMINATE result, final result
is given immediately as this result.}
\item {\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
\subsubsection{Request Structure}
{\selectlanguage{english}\upshape\color{black}
Request is made of RequestItem elements. Each RequestItem is evaluated
against Policy Rule and for each evaluation separate result is
generated as described above.RequestItem is made of 4 elements:}

{\color{black}
Subject - represents entity requesting specified action}

\liststyleWWviiiNumxix
\begin{itemize}
\item {\color{black}
Resource - destination/object of the action}
\item {\color{black}
Action - specifies what has to be done on resource}
\item {\color{black}
Context - for additional information which does not fit anywhere else,
like the current time.}
\end{itemize}
{\color{black}
Effectively RequestItem may have only one Subject, one Resource, one
Action and one Context. If there are more than one element of any kind
of sub-elelemt, then in the evaluator this RequestItem is split into
several items containing all possible permutations and results are
obtained for every item separately. How results are combined will be
explained later.}

{\color{black}
Additionally Subject could contain sub-elements SubjectAttribute. Those
are meant to represent different kinds of requesters{\textquotesingle}
identities. Example:}

\liststyleWWviiiNumxviii
\begin{itemize}
\item {\color{black}
Subject}

\begin{itemize}
\item {\color{black}
SubjectAttribute: name is ALICE}
\item {\color{black}
SubjectAttribute: age is YOUNG}
\item {\color{black}
SubjectAttribute: gender is GIRL}
\end{itemize}
\end{itemize}
{\color{black}
Context could also be made of ContextAttribute elements in the same way
as Subject.}

{\color{black}
The following is an example of the Request:}

{\itshape\color{black}
{\textless}Request
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\color{black}
\textit{\ }\textit{{\textless}RequestItem{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}Subject{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:x509:identity{\textquotedbl}{\textgreater}/O=KnowARC/OU=UiO/CN=Physicist{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ \ \ \ }\textit{{\textless}SubjectAttribute
AttributeId={\textquotedbl}urn:knowarc:voms:attribute{\textgreater}knowarc:atlasuser{\textless}/SubjectAttribute{\textgreater}}}

{\selectlanguage{english}\color{black}
\textit{\ \ \ }\textit{{\textless}/Subject{\textgreater}}}

>>>>

% This file was converted to LaTeX by Writer2LaTeX ver. 0.9.3

@@

\documentclass[a4paper]{article}

@@

\usepackage{amsmath}
\usepackage{amssymb,amsfonts,textcomp}

@@

\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=Aleksandr Konstantinov, pdfsubject=, pdfkeywords=}

@@

\renewcommand\thesection{\arabic{section}}
\renewcommand\thesubsection{\arabic{subsection}}
\renewcommand\thesubsubsection{\arabic{subsection}.\arabic{subsubsection}}
\renewcommand\theparagraph{\arabic{subsection}.\arabic{subsubsection}.\arabic{paragraph}}

@@

\newcommand\liststyleLii{%

@@

@@

\author{Aleksandr Konstantinov}
\date{2008-12-23}

@@

\bigskip
\clearpage\clearpage\pagestyle{Convertv}

@@

\liststyleWWviiiNumxiii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\textit{delegtion.pdp} -- Delegation PDP}
\end{itemize}
{\selectlanguage{english}\upshape\color{black}
The Delegation PDP is basically similar to Arc PDP, except it uses the
delegation policy parsed from remote proxy credential by
\textit{delegation.collector}, and evaluates the request against
delegation policy. See section \ref{bkm:Ref204009393}for the design
idea and use case of delegation policy in fine-grained identity
delegation.}

@@

{\selectlanguage{english}\itshape\color{black}
\ \ Rule (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Subjects (1)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ \ \ Attribute (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ Resources (0-1)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ Resource (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Actions (0-1)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ Action (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Conditions (0-1)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ Condition (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ \ \ Attribute (1-)}

@@

{\selectlanguage{english}\itshape\color{black}
\ \ RequestItem (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Subject (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ SubjectAttribute (1-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Resource (0-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Action (0-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ Context (0-)}

{\selectlanguage{english}\itshape\color{black}
\ \ \ \ \ \ ContextAttribute (1-)}

@@

\subsubsection{Basic Elements of Policy}
{\selectlanguage{english}\color{black}
There are 2 basic objects - {\textquotedbl}policy{\textquotedbl} and
{\textquotedbl}request{\textquotedbl}. There is 1 main actor -
Evaluator. Curretly there are two types of elements in policy: Policy
and Rule. Policy element is made of Rule elements.}

{\selectlanguage{english}\color{black}
Evaluator matches request to policy and produces one of 4 following
results:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\foreignlanguage{english}{\textbf{PERMIT}}\foreignlanguage{english}{ -
policy explicitely permits activity specified in request because
request matches some part of policy and corresponding effect specified
in policy is PERMIT. }}
\end{itemize}
{\selectlanguage{english}\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Rule:
}}}\foreignlanguage{english}{\textit{PERMIT person ALICE to PLAY in
place called WONDERLAND}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Request:
}}}\foreignlanguage{english}{\textit{person ALICE wants to PLAY in
place called WONDERLAND}}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\foreignlanguage{english}{\textbf{DENY}}\foreignlanguage{english}{ -
policy explicitely denies activity specified in Request because Request
matches some part of policy and corresponding effect specified in
policy is DENY.}}
\end{itemize}
{\selectlanguage{english}\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Rule: DENY
}}}\foreignlanguage{english}{\textit{fruit APPLE to GROW on PEACH
tree}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{\ \ \ \ \ \ Request}}}\foreignlanguage{english}{\textit{:
fruit APPLE to be GROWN on PEACH tree}}}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\foreignlanguage{english}{\textbf{INDETERMINATE}}\foreignlanguage{english}{
- request has some part which does not correspond to policy.}}
\end{itemize}
{\selectlanguage{english}\bfseries\itshape\color{black}
Example: }

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Rule}}}\foreignlanguage{english}{\textit{:
DENY fruit APPLE to GROW on PEACH tree}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
fruit APPLE to be GROWN on WHEAT ground}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
flower SUNFLOWER to be grown on PEACH tree}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{Explaination}}}\foreignlanguage{english}{:
Here, it is not possible to obtain any matching result - neither
positive (DENY or PERMIT) nor negative (NOT\_APPLICABLE, see below)}}

{\selectlanguage{english}\color{black}
In the request, the {\textquotedbl}ground{\textquotedbl} is completely
uncomparable to the {\textquotedbl}tree{\textquotedbl} in policy. One
can compare {\textquotedbl}PEACH tree{\textquotedbl} and
{\textquotedbl}APPLE tree{\textquotedbl} because they are both
{\textquotedbl}tree{\textquotedbl}; But it is impossible to compare
{\textquotedbl}PEACH tree{\textquotedbl} and {\textquotedbl}WHEAT
ground{\textquotedbl} becaue they of different kind (Policy is about
tree and Request is about ground).}

{\selectlanguage{english}\color{black}
In a similar way one can{\textquotesingle}t compare {\textquotedbl}fruit
APPLE{\textquotedbl} and {\textquotedbl}flower SUNFLOWER{\textquotedbl}
(here policy is about fruits and Request is about flower).}

{\selectlanguage{english}\color{black}
Any other situation which makes it impossibile to compare two attributes
will also cause {\textquotedbl}INDETERMINATE{\textquotedbl}.}

@@

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\foreignlanguage{english}{\textbf{NOT\_APPLICABLE}}\foreignlanguage{english}{
- all parts of the Request have corresponding parts in the Policy, but
some value of those parts are not the same. Hence request does not
match policy.}}
\end{itemize}
{\selectlanguage{english}\bfseries\itshape\color{black}
Example:}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Rule}}}\foreignlanguage{english}{\textit{:
DENY fruit APPLE to GROW on PEACH tree}}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
fruit APPLE to be GROWN on APPLE tree }}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
fruit ORANGE to be GROWN on PEACH tree }}}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textit{\ \ \ \ \ \ }}\foreignlanguage{english}{\textbf{\textit{Request}}}\foreignlanguage{english}{\textit{:
fruit ORANGE to be GROWN on APPLE tree }}}

{\selectlanguage{english}\color{black}
Explanation: for each part of the Request evaluator can find relevant
part in the Policy - both Policy and Request are about fruit and tree.
But the values do not match.}

{\selectlanguage{english}\color{black}
If it is required to reduce evaluation results to boolean value PERMIT
maps to TRUE and rest of results to FALSE.}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{\textit{Note}}}\foreignlanguage{english}{\textit{:
It would be useful to make it possible to specify secondary effect
which would become }}\foreignlanguage{english}{active in case Request
is NOT\_APPLICABLE. For example:}}

{\selectlanguage{english}\itshape\color{black}
DENY fruit APPLE to GROW on PEACH tree otherwise PERMIT}

{\selectlanguage{english}\color{black}
But one should be careful because example above would allow fruit PLUMS
to grow on APPLE trees :)}

{\selectlanguage{english}\color{black}
This kind of requirement can be supported by using the algorithm between
policies. \ For example, in case of above scenario, we can use some
algorithm like {\textquotedbl}Permit-if-notapplicable{\textquotedbl}.
See below the {\textquotedbl}Policy matching{\textquotedbl} part for
more explaination. \ }

\subsubsection{Policy MAtching}
{\selectlanguage{english}\color{black}
Policy is made of Rule elements. Request is evaluated against each Rule.
Each evaluation produces same results as policy evaluation described
above. The results from all Rules are then combined in order to produce
final result for whole policy. Results Combining Algorithm is specified
in Policy. There are 26 algorithms currently:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\color{black}
\foreignlanguage{english}{\textbf{\textit{Deny-Overrides}}}\foreignlanguage{english}{
- this is default if no algorithm specified.}}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If there is at least one DENY in results final result is DENY.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one PERMIT, the final result is PERMIT.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\bfseries\color{black}
NOTE: CHECK if algorithm description is correct}

{\selectlanguage{english}\color{black}
Special case is Policy wih no rules. Probably such policy shoudl be
treated as always producing DENY.}

{\selectlanguage{english}\bfseries\color{black}
Discussion}

{\selectlanguage{english}\itshape\color{black}
If there is no Rule under Policy, it means no restriction from this
Policy, what exactly shoud the Policy give? DENY or NOT\_APPLICABLE?}

{\selectlanguage{english}\itshape\color{black}
It seems to be more logical to produce INDETERMINATE because there is no
rule with elements which could be compared to request.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\bfseries\itshape\color{black}
Permit-Overrides.}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If there is at least one PERMIT in results final result is PERMIT.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\color{black}
Special case is Policy wih no rules. Probably such policy should be
treated as always producing DENY.}

{\selectlanguage{english}\bfseries\color{black}
NOTE: CHECK if algorithm description is correct}

{\selectlanguage{english}\bfseries\color{black}
Discussion}

{\selectlanguage{english}\itshape\color{black}
Question? The same as above.}

\liststyleWWviiiNumxxi
\begin{itemize}
\item {\selectlanguage{english}\bfseries\itshape\color{black}
Ordered algorithms.}
\end{itemize}
{\selectlanguage{english}\color{black}
These specify priorities for all four possible results. Their names look
like Result1-Result2-Result3-Result4 with Result\# naming result types,
for example Permit-Deny-NotApplicable-Indeterminate. The results are
combined in following way:}

\liststyleWWviiiNumxxi
\begin{itemize}
\item \begin{itemize}
\item {\selectlanguage{english}\color{black}
If there is at least one result of Result1 type then final result is
Result1.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one result of Result2 type then final
result is Result2.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one result of Result3 type then final
result is Result3.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is Result4.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\color{black}
There are 24 possible combinations of those algorithms.}

{\selectlanguage{english}\upshape\color{black}
\foreignlanguage{english}{\textbf{Note:}}\foreignlanguage{english}{ It
would be useful to have more combining algorithms. For example }}

@@

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If this Rule gives DENY in results, final result is DENY.}
\item {\selectlanguage{english}\color{black}
Otherwise if this Rule gives NOT\_APPLICABLE, final result is PERMIT.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is INDETERMINATE. }
\end{itemize}
\item {\selectlanguage{english}\color{black}
{\textquotedblleft}Permit-if-allPermit{\textquotedbl} - Permit if all
the Rules gives Permit, this algorithm is useful in case if we are
collecting different policies from a few sources, and we want the
request to satisfy all of them.}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If all of the Rule give PERMIT, the final result is PERMIT.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one DENY the final result is DENY.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is at least one NOT\_APPLICABLE final result is
NOT\_APPLICABLE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is INDETERMINATE.}
\end{itemize}
\item {\selectlanguage{english}\color{black}
OnlyOneApplicable}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If there is one gives INDETERMINATE, final result INDETERMINATE is given
immediately.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is exactly only one gives applicable result (DENY or
PERMIT), final result is as this result.}
\item {\selectlanguage{english}\color{black}
Otherwise if there is more than one gives applicable result, final
result is INDTERMINATE.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\color{black}
This algorithm makes sure that only one Rule is selected when making
decision.}

\liststyleWWviiiNumxvii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
FirstApplicable}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
If there is one give DENY, PERMIT or INDETERMINATE result, final result
is given immediately as this result.}
\item {\selectlanguage{english}\color{black}
Otherwise final result is NOT\_APPLICABLE.}
\end{itemize}
\end{itemize}
\subsubsection{Request Structure}
{\selectlanguage{english}\color{black}
Request is made of RequestItem elements. Each RequestItem is evaluated
against Policy Rule and for each evaluation separate result is
generated as described above.RequestItem is made of 4 elements:}

{\selectlanguage{english}\color{black}
Subject - represents entity requesting specified action}

\liststyleWWviiiNumxix
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Resource - destination/object of the action}
\item {\selectlanguage{english}\color{black}
Action - specifies what has to be done on resource}
\item {\selectlanguage{english}\color{black}
Context - for additional information which does not fit anywhere else,
like the current time.}
\end{itemize}
{\selectlanguage{english}\color{black}
Effectively RequestItem may have only one Subject, one Resource, one
Action and one Context. If there are more than one element of any kind
of sub-elelemt, then in the evaluator this RequestItem is split into
several items containing all possible permutations and results are
obtained for every item separately. How results are combined will be
explained later.}

{\selectlanguage{english}\color{black}
Additionally Subject could contain sub-elements SubjectAttribute. Those
are meant to represent different kinds of requesters{\textquotesingle}
identities. Example:}

\liststyleWWviiiNumxviii
\begin{itemize}
\item {\selectlanguage{english}\color{black}
Subject}

\begin{itemize}
\item {\selectlanguage{english}\color{black}
SubjectAttribute: name is ALICE}
\item {\selectlanguage{english}\color{black}
SubjectAttribute: age is YOUNG}
\item {\selectlanguage{english}\color{black}
SubjectAttribute: gender is GIRL}
\end{itemize}
\end{itemize}
{\selectlanguage{english}\color{black}
Context could also be made of ContextAttribute elements in the same way
as Subject.}

{\selectlanguage{english}\color{black}
The following is an example of the Request:}

{\selectlanguage{english}\itshape\color{black}
{\textless}Request
xmlns={\textquotedbl}http://www.nordugrid.org/schemas/request-arc{\textquotedbl}{\textgreater}}

{\selectlanguage{english}\itshape\color{black}
\ {\textless}RequestItem{\textgreater}}

{\selectlanguage{english}\itshape\color{black}
\ \ \ {\textless}Subject{\textgreater}}
